---
title: "IO"
author: "Ignacio Vellido"
date: "10/28/2020"
output:
  html_document:
    theme: darkly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results="hold")
library(tidyverse)
```

---

Ask the user to introduce with the kwyboard a string s and a number n and show in the screen n times the string s (without spaces between words) / Pide al usuario que introduzca por teclado un string s y un número n y que muestre en pantalla n veces seguidas el string s (sin espacios entre palabra y palabra).
s="hola, n=3
holaholahola

```{r}
#s <- scan("", what=character())
#n <- scan("")
s <- "hola"
n <- 3
paste(rep(s, n), sep="", collapse="")
```

---

Create three files called dos.txt, tres.txt and cinco.txt containing the table of 2, 3 and 5 respectively (the first 10 values of each, one number at a time) / Crea tres ficheros llamados dos.txt, tres.txt y cinco.txt que contenga la tabla del 2, la del 3 y la del 5 respectivamente (los primeros 10 valores de cada tabla, un número en cada línea de cada fichero).

```{r}
write.table(matrix(1:10 * 2), "dos.txt", row.names=F, col.names=F)
print(read.table("./dos.txt"))
```


```{r}
write.table(matrix(1:10 * 3), "tres.txt", row.names=F, col.names=F)
print(read.table("./tres.txt"))
```


```{r}
write.table(matrix(1:10 * 5), "cinco.txt", row.names=F, col.names=F)
print(read.table("./cinco.txt"))
```

---

Load the three files created in the last exercise and create a matrix with three columns, one for each file conent / Carga los tres ficheros creados en el punto anterior y construye una matriz que, en cada columna, tengo el contenido de cada fichero.

```{r}
dos <- read.table("dos.txt")
tres <- read.table("tres.txt")
cinco <- read.table("cinco.txt")
m <- bind_cols(dos,tres,cinco)
m
```

---

Write the first five rows of the matrix created in the last exercise in a new file called prime.txt and the last five in another file called fin.txt. Both files must have the data separated by commas / Escribe las cinco primera filas de matriz del ejercicio anterior en un fichero nuevo llamado prime.txt y las cinco últimas en otro fichero llamado fin.txt. Ambos ficheros deben tener los datos separados por comas.

```{r}
write.table(m[1:5,1:3], "prime.txt", sep=",", row.names=F, col.names=F)
print(read.table("./prime.txt"))
```


```{r}
write.table(m[6:10,1:3], "fin.txt", sep=",", row.names=F, col.names=F)
print(read.table("./fin.txt"))
```

---

Given two numbers, f and c, introduced using the keyboard by the user, create a square of f rows and c collumns with character "x". See an example for f=4 and c=3 below: / Dados dos números introducidos por el usuario f y c, crea un cuadrado de f filas y c columnas con el caracter "x". Un ejemplo con f=4 y c=3 sería:

xxx

xxx

xxx

xxx

```{r}
#f <- scan("")
#c <- scan("")
f <- 4
c <- 3

matrix(rep("x", f*c), nrow = 4)
```